### Newer version in Y:\Macro\Visualisation\Choropleth\docs\EUROSTAT

Attention please!
=============

Forest sector employment in the European Union
========================================================
This scripts loads EUROSTAT forestry datasets and displays data.

Original datasets and metadata can be viewed on the [eurostat website](http://epp.eurostat.ec.europa.eu/portal/page/portal/forestry/data/database). See also the Eurostat working group on [forestry statistics and accounts](https://circabc.europa.eu/w/browse/8be55496-f9fe-4c00-923e-bbebf52f0ec6).

There was a change in classification in 2008. Therefore EUROSTAT has 2 employment tables for the forest sector: "for_emp_lfs1" Before 2008 and "for_emp_lfs" after 2008. We pull data from those 2 tables and mix them in some graph. Be aware that the employment figures before and after 2008 __are not comparable__. The difference by country for the year 2008 is computed in a table below.

Table of content
----------------
* [Load data from EUROSTAT](#load)
* [Clean data](#clean)
* Plots
 * [Employment by subsector in EU28](#empEU28)
 * [Employment by subsector by country](#empCountry)

```{r echo=FALSE}
# Set same path for knitr evaluation as for interactive use
opts_knit$set(root.dir = '../..')
```

<a name="load"></a>
Load data from EUROSTAT
---------------------------------------
```{r, eval=FALSE}
library(SmarterPoland)
for_emp_lfs = getEurostatRCV(kod = "for_emp_lfs")
```
The above code chunk is not evaluated.
The current script loads the dataset from a local file.
(Because it's a few seconds faster). No difference should be seen.

```{r }
# Load local dataset on computer. Skip this if you use getEurostatRCV() above.
load("rawdata/EUROSTAT_forestry.RDATA")
library(ggplot2)
library(xtable)
R.version.string
```

Description of the forestry employment __for_emp_lfs__ dataset :
* __unit__:
* __sex__: female (F), male (M) or total (T)
* __isced97__: education level
* __wstatus__, working status: employed persons (EMP), employees (SAL), self employed persons (SELF)
* __nace_r2__, sub sector of employment, forestry, manufacture of wood products or manufacture of paper products (see also below)
* __geo__: geographical area
* __time__: year
* __value__: number of employed persons in thousand

Sub-sectors of employment in forestry datasets:
* Before 2008, in for_emp_lfs1 : `r unique(for_emp_lfs1$nace_r1)`
* After 2008, in for_emp_lfs : `r unique(for_emp_lfs$nace_r2)`

<a name="clean"></a>
### Clean data
```{r}
# Change time to a numeric value
for_emp_lfs1$time = as.numeric(as.character(for_emp_lfs1$time))
for_emp_lfs$time = as.numeric(as.character(for_emp_lfs$time))

# Add a readable description of subsector of employment.
# Rename column and factor values. 
# levels() keeps these columns as factor (The use of factors may not be needed).
for_emp_lfs1$subsector = for_emp_lfs1$nace_r1
levels(for_emp_lfs1$subsector)[levels(for_emp_lfs1$nace_r1)=="A0202"] = "Forestry and logging"
levels(for_emp_lfs1$subsector)[levels(for_emp_lfs1$nace_r1)=="DD20"] = 
    "Manufacture of wood products except furniture"
levels(for_emp_lfs1$subsector)[levels(for_emp_lfs1$nace_r1)=="DE21"] = "Manufacture of paper products"

# Same manipulation for data after 2008
for_emp_lfs$subsector = for_emp_lfs$nace_r2
levels(for_emp_lfs$subsector)[levels(for_emp_lfs$nace_r2)=="A02"] = "Forestry and logging"
levels(for_emp_lfs$subsector)[levels(for_emp_lfs$nace_r2)=="C16"] = 
    "Manufacture of wood products except furniture"
levels(for_emp_lfs$subsector)[levels(for_emp_lfs$nace_r2)=="C17"] = "Manufacture of paper products"
```


Plots 
-----
<a name="empEU28"></a>
### Employment by subsector in EU28
There is no aggregated EU data before 2001. See below for country level data from 1992.

Time series before 2008 with subsector classification nace_r1
```{r fig.width=12}
# Years before 2001 are empty
EU28_r1 = subset(for_emp_lfs1, sex=="T" & isced97=="TOTAL" & wstatus=="EMP" & geo=="EU28" & time>2001 )
p = ggplot() +
    geom_point(data=EU28_r1, aes(x=time,y=value, color=subsector)) +
    ylab("Thousand employees") 
# plot(p) # Uncomment to plot only before 2008
```

Time series after 2008 with subsector classification nace_r2
```{r fig.width=12}
EU28 = subset(for_emp_lfs, sex=="T" & isced97=="TOTAL" & wstatus=="EMP" & geo=="EU28" )
p + geom_point(data=EU28, aes(x=time,y=value, color=subsector)) 
```


<a name="empCountry"></a>
### Employment by subsector by country
Time series before 2008
```{r fig.width=12}
emp1 = subset(for_emp_lfs1, sex=="T" & isced97=="TOTAL" & wstatus=="EMP" & !is.na(value) &
                 !geo %in% c("EA13", "EA17", "EA18", "EU15", "EU27", "EU28"))
p =  ggplot() + geom_line(data=emp1, aes(x=time,y=value, color=subsector)) +
    ylab("Thousand employees") +
    scale_x_continuous(breaks=seq(1990,2010,10), limits=c(1990, 2013))
# plot(p + facet_wrap(~geo)) # Uncomment to plot only before 2008
```

Time series before and after 2008
```{r warning=FALSE, fig.width=12}
# Select countries only
emp = subset(for_emp_lfs, sex=="T" & isced97=="TOTAL" & wstatus=="EMP" & !is.na(value) &
                 !geo %in% c("EA13", "EA17", "EA18", "EU15", "EU27", "EU28"))
p = p + geom_line(data=emp, aes(x=time,y=value, color=subsector)) 
```

Plot employment by country with a fixed scale
```{r fig.width=12}
plot(p + facet_wrap(~geo))
```

Plot employment by country with a free scale
```{r fig.width=12}
plot(p + facet_wrap(~geo, scales = "free_y"))
```

### Compare 2008 values between first and second classification.
```{r, results='asis'}
emp1_8 = emp1[emp1$time==2008,c("subsector", "geo", "value")]
emp$valueNewClassif = emp$value
emp_8 = emp[emp$time==2008,c("subsector", "geo", "valueNewClassif")]
empComp = merge(emp1_8,emp_8)

empComp$diff = empComp$value - empComp$valueNewClassif
empComp = transform(empComp,
                    diff = valueNewClassif - value, 
                    diffPercent = round((valueNewClassif - value)/value*100))
print(xtable(empComp), row.names=FALSE, type = "html")
```

### Map of forest sector employment per 1000 ha of forest



Future visualisations
---------------------
* Map of forest sector employment per 1000 m^3
* Employment by type (employee or self employed) and by sector
