Demand for Sawnwood in the European Union
=========================================

```{r setup, include=FALSE}
library(ggplot2)
opts_chunk$set(cache=TRUE) # Set global chunk options
options(width=80)
```


### Load data
FAOSTAT is the source of forest products data and
World Bank is the source of GDP, deflator and exchange rate data.  
```{r load, results='show', warning=FALSE, message=FALSE}
setwd("../..")
print(load("enddata/EU27 sawnwood demand.rdata"))
EU = read.csv("rawdata/EUCountries.csv")
swd = sawnwood # Give a shorter name to the data frame
```

Consumption, Production and Trade Volumes
-----------------
In the following file, we explore a data table containing paper and paperboard consumption data for __`r length(unique(swd$Country))`__ countries from `r min(swd$Year)` to `r max(swd$Year)`.  
__In `r max(swd$Year)`, the overall EU consumption, production and trade in million Tons per item was:__
```{r, echo=FALSE}
swdaggLast = subset(swdagg, Year==max(swdagg$Year)&
                       !Element%in%c("Import_Value", "Export_Value","Price"))
swdaggLast$Value = round(swdaggLast$Value/1e6)
swdaggLast = reshape(swdaggLast, drop=c("Year", "Deflator"),
                idvar="Item", timevar="Element",direction="wide")

# names(swdaggLast) = sub("Value.","",names(swdaggLast)) 
names(swdaggLast) = c("Item", "Consumption", "Production",
                     "Net_Trade", "Import", "Export")
print(swdaggLast) # This will be displaid on the HTML page
```



Consumption, production, trade and net trade of Sawnwood in the European Union
```{r consumptionTotalEU, fig.width=13, fig.height=6}
ggplot(data=subset(swdagg, !Element%in%c("Import_Value", "Export_Value","Price"))) +  
    geom_line(aes(x=Year, y=Value/1e6, colour=Item),size=1) +
    facet_wrap(~Element) +
    ylab("Million metric Ton") + theme_bw()
 
```


Relationship between log(Consumption) and log(GDP) 
--------------------------------------------------
### Total Sawnwood
Data for 2000 and beyond in red
Data for EU15 countries in Blue
```{r}
plot(log(Consumption) ~ log(GDPconstantUSD), data=subset(swd, Item=="Total Sawnwood"))
points(log(Consumption) ~ log(GDPconstantUSD), 
     data=subset(swd, Item=="Total Sawnwood"&Year>1999), col="red")
# swd = merge(swd, head(subset(EU, select=c(Country,EU15))))
```


### Explore the influence of Year and Net_Trade by country
```{r warning=FALSE, fig.width=12, fig.height=8}

p = ggplot(data=subset(swd, Item!="Total Sawnwood"),
           aes(x=log(GDPconstantUSD), y=log(Consumption))) + facet_wrap(~Item)
p + scale_shape_identity() +
    geom_point(aes(color=Country, shape=1 + 2*(Net_Trade>0) ))
```


### Plot for Total Sawnwood 
```{r warning=FALSE, fig.width=12, fig.height=8}

p = ggplot(data=subset(swd, Item=="Total Sawnwood"),
           aes(x=log(GDPconstantUSD), y=log(Consumption))) + facet_wrap(~Item)
p + scale_shape_identity() +
    geom_point(aes(color=Country, shape=1 + 2*(Net_Trade>0) ))
```

# Sort countries by net trade - Maybe this can be put in the clean script?
