Demand for paper and paperboard products in the European Union
========================================================

```{r setup, include=FALSE, cache=FALSE}
library(ggplot2)
library(plm)
library(plyr)
opts_knit$set(root.dir = '../..') # Change to project root directory
opts_chunk$set(cache=TRUE) # Set global chunk options
options(width=80)
```

### Load data
FAOSTAT is the source of forest products data and
World Bank is the source of GDP, deflator and exchange rate data.  
```{r load, results='show', warning=FALSE, message=FALSE, cache=FALSE}
source("code/func.r")
print(load("enddata/EU27 paper products demand.rdata"))
pp = paperProducts # Give a shorter name to the data frame
```

<a name="CPT"></a>

Consumption, Production and Trade Volumes
-----------------
In the following file, we explore a data table containing paper and paperboard consumption data for __`r length(unique(pp$Country))`__ countries from `r min(pp$Year)` to `r max(pp$Year)`.  
__In `r max(pp$Year)`, the overall EU consumption, production and trade in million Tons per item was:__
```{r, echo=FALSE}
ppaggLast = subset(ppagg, Year==max(ppagg$Year)&
                       !Element%in%c("Import_Value", "Export_Value","Price"))
ppaggLast$Value = round(ppaggLast$Value/1e6)
ppaggLast = reshape(ppaggLast, drop=c("Year", "Deflator"),
                idvar="Item", timevar="Element",direction="wide")
# names(ppaggLast) = sub("Value.","",names(ppaggLast)) 
names(ppaggLast) = c("Item", "Consumption", "Production",
                     "Net_Trade", "Import", "Export")
print(ppaggLast) # This will be displaid on the HTML page

# Prepare a table of big consumers and big exporters for later use
bigconsumers = subset(pp, Year==2012&Consumption>5e6&Item=="Total Paper and Paperboard", 
                     select=c("Year","Country","Consumption"))
bigexporters  = subset(pp, Year==2012&Net_Trade>5e6&Item=="Total Paper and Paperboard", 
                     select=c("Year","Country","Net_Trade"))
```


```{r consumptionTotalEU, fig.width=13, fig.height=6}
# Consumption, production, trade and net trade at the European Level
ggplot(data=subset(ppagg, !Element%in%c("Import_Value", "Export_Value","Price"))) +  
    geom_line(aes(x=Year, y=Value/1e6, colour=Item),size=1) +
    facet_wrap(~Element) +
    ylab("Million metric Ton") + theme_bw()
 
```
Net trade is slightly in favor of exports for most products except for newsprint. From bilateral trade statistics (not shown here), we know that most of the trade is happening inside the European Union.


<a name="Consumption"></a>
Consumption at the country level
--------------------------------
<a name="TCbyCountry"></a>
### Consumption by country 
Apparent consumption of Total paper and paperboard by country 
```{r consumptionbyCountry, fig.width=13, fig.height=8}
ggplot(data=subset(pp, Item=="Total Paper and Paperboard"))+
    geom_line(aes(x=Year, y=Consumption/1e+06, colour=Item), size=1) + 
    facet_wrap(~Country, scales = "free_y") +
    xlab("Year") + ylab("Million metric Ton") + 
    theme_bw() +  theme(legend.position="bottom")
```
The largest consumers of paper and paperboard products are `r bigconsumers$Country`.


<a name="CGPOpapbyCountry"></a>

### Apparent consumption of graphics paper and other paper and paperboard by country 
```{r graphAndOtherconsumptionbyCountry, fig.width=13, fig.height=8}
# Create a new table with graphics papers following Lauri's comments
ppgraph = subset(pp, Item %in% c("Newsprint","Printing and Writing Paper"))
ppgraph = aggregate(ppgraph[c("Consumption", "Net_Trade")],
                     ppgraph[c("Year", "Country")], sum)
ppgraph$Item = "Graphics Paper"
ppgraph = rbind(ppgraph, 
                subset(pp, Item == "Other Paper and Paperboard", 
                       c(Year, Country, Item, Consumption, Net_Trade)))

ggplot(data=ppgraph)+
    geom_line(aes(x=Year, y=Consumption/1e+06, colour=Item), size=1) + 
    facet_wrap(~Country, scales = "free_y") +
    xlab("Year") + ylab("Million metric Ton") + 
# scale_colour_manual(values = c("Blue", "Green")) +
    theme_bw() +  theme(legend.position="bottom") 

```

### Heat map of consumption per capita
Consumption of total paper and paperboard in $$m^3$$ per habitant.
```{r heatmap}
# add population data
pp = merge(pp, wb[c("Country","Population")])
ggplot(data=subset(pp, Item=="Total Paper and Paperboard"), aes(Year, Country)) +
    geom_tile(aes(fill = 1000*Consumption/Population), colour = "white") +
    scale_fill_gradient(low = "white", high = "steelblue")

ggplot(data=pp, aes(Year, Country)) +
    geom_tile(aes(fill = 1000*Consumption/Population), colour = "white") +
    scale_fill_gradient(low = "white", high = "steelblue") + facet_wrap(~Item)
```


<a name="CNTbyCountry"></a>
### Consumption and Net Trade by Country
Apparent consumption and Net Trade of Total paper and paperboard by country 
```{r consumptionNetTradebyCountry, fig.width=13, fig.height=8}
ggplot(data=subset(pp, Item=="Total Paper and Paperboard"))+
    geom_line(aes(x=Year, y=Consumption/1e+06, colour=Item, linetype = "Consumption"), 
              size=1) + 
    geom_line(aes(x=Year, y=Net_Trade/1e+06, colour=Item, linetype = "Net trade"),
              size=1) + 
    facet_wrap(~Country, scales = "free_y") +
    xlab("Year") + ylab("Million metric Ton") + 
    theme_bw() +  theme(legend.position="bottom")
```
Large producing countries such as Finland and Sweden produce over 10 million tons of paper products per year and export most of their production. Germany is the largest consumer with over 20 million tons of paper products consumed annually, Germany's net trade is slightly positive. France is the second largest consumer and is slightly importing. The United Kingdom also consumes around 10 million tons annually and is importing more than half of its production.

<a name="GDPbyCountry"></a>

### GDP by country
```{r GDPbyCountry, fig.width=13, fig.height=8, warning=FALSE}
# 1 plot per country
ggplot(data=wb, aes(x=Year, y=GDPconstantUSD/1e9)) +  
    geom_line(size=1) + xlab("Years") + ylab("GDP (constant) in 2010 Billion USD ") +
    theme_bw() + theme(legend.position= "none") + facet_wrap(~Country, scales = "free_y")

```


<a name="Prices"></a>

Prices of paper and paperboard
----------------
### EU prices
Average trade price of paper products in the European Union
```{r priceEU, fig.width=10, fig.height=6}
ggplot(data=subset(ppagg, Element=="Price")) +
    geom_line(aes(x=Year, y=Value, color=Item)) +
    ylab("Price in (2010) constant USD /Metric Ton")
```
Price has lowered since 2010, probably due to overcapacity in the market.

### Prices per country
Prices are a ponderation of import and export prices as used in Chas-Amil and Buongiorno 2000. Prices are expressed in constant US dollars of 2010.
```
Price = (Import_Value + Export_Value)/ (Import_Quantity + Export_Quantity) / DeflUS *1000)
```
Countries with a particularly high price were removed from the set. 
```{r pricesByCountry, fig.width=12, fig.height=8, warning=FALSE}
library(scales)
ggplot(data=subset(pp, !Country%in%c("Slovakia", "Malta"))) +
    geom_line(aes(x=Year, y=Price, color=Item))  + facet_wrap( ~ Country) + 
    ylab("Price in (2010) constant USD /m3") + scale_y_continuous(labels = dollar)

```

### Trade prices by Country
```{r tradePricesByCountry, fig.width=12, fig.height=8, warning=FALSE}
ggplot(data=subset(pptrade, !Country%in%c("Slovakia", "Malta"))) +
    geom_line(aes(x=Year, y=Price_Trade, color=Item, linetype = Trade))+
    facet_wrap( ~ Country) + 
    ylab("Price in (2010) constant USD /m3")
```

### Trade prices for Sweden, Germany, France and Finland
```{r tradePricesSGFF, fig.width=12, fig.height=8, warning=FALSE}
# Sweden Germany France Finland 
ggplot(data=subset(pptrade, Country%in%c("Sweden", "Germany", "France", "Finland"))) +
    geom_line(aes(x=Year, y=Price_Trade, color=Item, linetype = Trade))  + 
    facet_wrap( ~ Country) + ylab("Price in (2010) constant USD /m3")
```

<a name="Distribution"></a>

Distribution of the estimation data for 4 important years
---------------------------------------------------------
## Cross sectional distribution of demand 
Number of countries for a given demand volume in million m3
```{r histdemand, fig.width=12, fig.height=8, warning=FALSE}
# 1980, 1990, 2000, 2010
pp4hist = subset(pp, Year%in%c(1980, 1990, 2000, 2010)&Item=="Total Paper and Paperboard")
ggplot(data=pp4hist) + geom_histogram(aes(x=Consumption/1e6), binwidth=1) + facet_wrap(~Year)
```
Number of countries for a given demand volume in log of the volume in m3
```{r histdemandlog, fig.width=12, fig.height=8, warning=FALSE}
# Consumption in log
ggplot(data=pp4hist) + geom_histogram(aes(x=log(Consumption)), binwidth=1) + facet_wrap(~Year)
```

## Distribution of prices
```{r histGDPlog1, fig.width=12, fig.height=8, warning=FALSE}
ggplot(data=pp4hist) + geom_histogram(aes(x=log(Price)), binwidth=0.1) + facet_wrap(~Year)
```

## Distribution of GDP
```{r histGDPlog2, fig.width=12, fig.height=8, warning=FALSE}
ggplot(data=pp4hist) + geom_histogram(aes(x=log(GDPconstantUSD)), binwidth=1) + facet_wrap(~Year)
```


<a name="EstimationData"></a>
Plot log of the estimation data 
-------------------------------
We will estimate the model
$$ log(Consumption) = \beta_0 + \beta_1 log(GDP) + \beta_2 log(Price) +
\beta_3 log(Consumption_{t-1}) $$

Lets look at the relationship between log(Consumption) and log(GDP) first.
### Total Paper and Paperboard
Data for 2000 and beyond in red
```{r}
plot(log(Consumption) ~ log(GDPconstantUSD), data=subset(pp, Item=="Total Paper and Paperboard"))
points(log(Consumption) ~ log(GDPconstantUSD), 
     data=subset(pp, Item=="Total Paper and Paperboard"&Year>1999), col="red")
```

### Explore the influence of Year and Net_Trade by country
Sort countries by Net_Trade, then display a color for each country
```{r warning=FALSE, fig.width=12, fig.height=8}
d=data.frame(s=c(45,43))
ppc = subset(pp, Item=="Total Paper and Paperboard"&Year==2012)
ppc = ppc[order(ppc$Net_Trade),]
pp$Country = factor(pp$Country, levels= ppc$Country, ordered=TRUE)
p = ggplot(pp, aes(x=log(GDPconstantUSD), y=log(Consumption))) + facet_wrap(~Item)
p + geom_point(aes(alpha=Year, color=Country))  
```

Check the [ggplot shape parameter](http://sape.inf.usi.ch/quick-reference/ggplot2/shape) and the  `scale_shape_identity` to improve this graph with a "+" sign for positive net trade and a "o" sign for negative net trade. Plot only Printing and Writing Paper for this graph.
```{r warning=FALSE, fig.width=12, fig.height=8}

p = ggplot(data=subset(pp, Item=="Printing and Writing Paper"),
           aes(x=log(GDPconstantUSD), y=log(Consumption))) + facet_wrap(~Item)
p + scale_shape_identity() +
    geom_point(aes(color=Country, shape=1 + 2*(Net_Trade>0) ))
```


### Total Paper and Paperboard
```{r}
plot(log(Consumption) ~ log(Price), data=subset(pp, Item=="Total Paper and Paperboard"))
points(log(Consumption) ~ log(Price), 
       data=subset(pp, Item=="Total Paper and Paperboard"&Year==2012), col="red")
```

### All products
```{r, warning=FALSE, fig.width=12, fig.height=8}
p = ggplot(pp,aes(x=log(Price), y=log(Consumption), alpha=Year)) + facet_wrap(~Item)
p + geom_point(aes(color=Country, alpha=Year)) 
```
Countries that trade less seem to have higher prices.

Change point shapes with a "+" sign for positive net trade and a "o" sign for negative net trade. Show only Printing and writing paper for this graph.
```{r, warning=FALSE, fig.width=12, fig.height=8}
p = ggplot(data=subset(pp, Item=="Printing and Writing Paper"),
           aes(x=log(Price), y=log(Consumption))) + facet_wrap(~Item)
p +  scale_shape_identity() + 
    geom_point(aes(color=Country, alpha=Year, shape=1 + 2*(Net_Trade>0) )) 
```

<a name="panelModel"></a>
Panel data models
-----------------
Over the 1960 - 1992 estimation period.
```{r}
EU15 = unique(wb$Country[wb$EU15==1])

# Create a column consumption at year t-1
pp = ddply(pp, .(Country,Item), mutate, 
           Consum_t_1 = c(NA,Consumption[-length(Consumption)]))

# Select same data as in the chasamil paper
chasamil = subset(pp, Year>=1969&Year<=1992&Country%in%EU15 & 
                  Country!="Ireland" & #Remove Ireland because it doesn't have GDPConstantUSD Data
                      Item=="Printing and Writing Paper", select=-Item)

# Set data as panel data
pchasamil <- plm.data(chasamil, index=c("Country","Year"))


```

### Pooled OLS
```{r, results='asis'}
pooling <- plm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1), 
               data=pchasamil, model= "pooling")
printSummaryTable(pooling)
```
```{r}
summary(pooling)
```

### Fixed effects or within estimator
```{r, results='asis'}
fixed <- plm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1), 
               data=pchasamil, model= "within")
printSummaryTable(fixed)
```

### Random effects estimator
```{r, results='asis'}
random <- plm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1), 
               data=pchasamil, model= "random")
printSummaryTable(random)
```
The estimation of the random effect model tells me that there is an Error in swar(object, data, effect) : "the estimated variance of the individual effect is negative".


