Paper Product Consumption models
========================================================

```{r setup, include=FALSE, cache=FALSE}
library(ggplot2)
library(plm)
library(plyr)
library(reshape2)
opts_knit$set(root.dir = '../..') # Change to project root directory
#opts_chunk$set(cache=TRUE) # Set global chunk options
options(width=80)
```


Data
----
### Load 
FAOSTAT is the source of forest products data and
World Bank is the source of GDP, deflator and exchange rate data.  
```{r load, results='show', warning=FALSE, message=FALSE, cache=FALSE}
source("code/func.r")
print(load("enddata/EU27 paper products demand.rdata"))
pp <- paperProducts # Give a shorter name to the data frame
EU15 <- unique(wb$Country[wb$EU15==1])
```

### Prepare data
```{r prepare}
# Create a column consumption at year t-1
pp <- ddply(pp, .(Country,Item), mutate, 
           Consum_t_1 = c(NA,Consumption[-length(Consumption)]))

# Select same data as in the chasamil paper
chasamil <- subset(pp, Year>=1969&Year<=1992&Country%in%EU15 & 
                      Item=="Printing and Writing Paper", select=-Item)

```
Le panel est il cylindré?

OLS by country
--------------
The Worldbank databank doesn't contain a deflator for Ireland over the time period of interest. Therefore we couldn't calculate a GDP in constant USD for Ireland.

### Serial temporal autocorrelation on one country
```{r}
# Change to a time series
consfr <- subset(chasamil,Country=="France")
consfr.ts <- ts(consfr$Consumption, start=min(consfr$Year))
plot(consfr.ts)
acf(consfr.ts,lag.max=100)
# numerical values of this autocorrelation function
a <- acf(consfr.ts,lag.max=100,plot=FALSE)
class(a)


# Autocorrelation coefficients to be applied on all countries
autocorr <- function(dtf){
    cons.ts <- ts(dtf$Consumption, start=min(dtf$Year))
    return(acf(cons.ts,lag.max=100,plot=FALSE)$acf)
}
a = ddply(chasamil, .(Country), autocorr)
autocorrTable <- melt(a, id.vars = c("Country"))
ggplot(autocorrTable) + aes(x=variable, y=value) + geom_bar() + facet_wrap(~Country)
```


### For all countries in the dataset
```{r }
#Remove Ireland because it doesn't have GDPConstantUSD Data
chasamil <- subset(chasamil, Country!="Ireland")
# Remove 1969 for Germany because Germany doesn't have GDP for that year
chasamil <- subset(chasamil, !(Country=="Germany"&Year==1969))

# A function that applies the dynamic model to a country 
dynModelCountry = function(dtf){
  dym = lm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1),
           data=dtf)
  coefs = data.frame(t(summary(dym)$coefficients[,1:2]))
  coefs$Country = unique(dtf$Country)
  if (nrow(coefs)==2){
    coefs$R_Squared  = c(summary(dym)$r.squared,NA)
  }
  return(coefs)
}

# Test the model for one country
dynModelCountry(subset(chasamil,Country=="France"))

# # Apply the dynamic model to each and all countries
x = ddply(chasamil, .(Country), dynModelCountry)

# Reorder and rename columns for esthetical reasons
x = x[c( "Country", "log.GDPconstantUSD.", "log.Price.", 
         "log.Consum_t_1.", "R_Squared")]
names(x) = c("Country", "Y", "P", "Dt_1", "R2")
x$Country[is.na(x$R2)] = ""
x[,-1] = round(x[,-1],2)
x$Y[is.na(x$R2)] = paste("(", x$Y[is.na(x$R2)], ")", sep="")
x$P[is.na(x$R2)] = paste("(", x$P[is.na(x$R2)], ")", sep="")
x$Dt_1[is.na(x$R2)] = paste("(", x$Dt_1[is.na(x$R2)], ")", sep="")

# Print table to html
print(xtable(x, caption="Demand equations for total paper and paperboard by country", 
             label="DemandByCountry"),  type = "html")

```

On the subject of Cochrane Orcutt method, [John Fox wrote](https://stat.ethz.ch/pipermail/r-help/2002-January/017774.html): "I'm not sure why you'd want to use these methods, other than for 
historical reasons." And he gave a function cochrane.orcutt.lm which takes a linear-model object as an input.


Panel data models
-----------------
Over the 1960 - 1992 estimation period.
```{r}

# Set data as panel data
pchasamil <- plm.data(chasamil, index=c("Country","Year"))
```

### Pooled OLS
```{r, results='asis'}
pooling <- plm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1), 
               data=pchasamil, model= "pooling")
printSummaryTable(pooling)
```
```{r}
summary(pooling)
```

### Fixed effects or within estimator
```{r, results='asis'}
fixed <- plm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1), 
               data=pchasamil, model= "within")
printSummaryTable(fixed)
```

### Random effects estimator
```{r, results='asis'}
random <- plm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1), 
               data=pchasamil, model= "random")
printSummaryTable(random)
```
The estimation of the random effect model tells me that there is an Error in swar(object, data, effect) : "the estimated variance of the individual effect is negative".

