Paper Product Consumption models
========================================================
* change prices to fictive euro prices to see if exchange rates 
have an effect on the significance of the estimated price elasticities.

```{r setup, include=FALSE, cache=FALSE}
library(ggplot2)
library(plm)
library(plyr)
library(reshape2)
opts_knit$set(root.dir = '../..') # Change to project root directory
#opts_chunk$set(cache=TRUE) # Set global chunk options
options(width=80)
```


Load Data
----
FAOSTAT is the source of forest products data and
World Bank is the source of GDP, deflator and exchange rate data.  
```{r load, results='show', warning=FALSE, message=FALSE, cache=FALSE}
source("code/func.r")
load("enddata/EU28 paper products base year 1987.rdata")
load("enddata/world Bank GDP defl pop 1987.rdata")
pp <- paperproducts$entity # Give a shorter name to the data frame
# Add GDPconstantEUR
pp <- merge(pp, wb[c("Country", "Year", "GDPconstantEUR")], all.x=TRUE)
EU15 <- unique(wb$Country[wb$EU15==1])
```

```{r prepare_data}
# Create a column consumption at year t-1
pp <- ddply(pp, .(Country,Item), mutate, 
           Consum_t_1 = c(NA,Consumption[-length(Consumption)]))

# Select same data as in the chasamil paper
chasamil <- subset(pp, Year>=1969&Year<=1992&Country%in%EU15)
# Remove 1969 for Germany because Germany doesn't have GDP for that year
chasamil <- subset(chasamil, !(Country=="Germany"&Year==1969))

# Subset only total value for the first estimations on country time series
chasamil_tot <- subset(chasamil, Item=="Total Paper and Paperboard")
```

Le panel est il cylindré?

OLS by country
--------------
### For all countries in the dataset
```{r }

# A function that returns coefficients and sd of the dynamic model to a country 
dynModelCountry = function(dtf){
  dym = lm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1),
           data=dtf)
  dym$Country = unique(dtf$Country)
  return(dym)
}

# Model for one country 
dynm_fr <- dynModelCountry(subset(chasamil_tot,Country=="France"))
country_series_table(dynm_fr)
# Othere way to display it 
print(xtable(dynm_fr), type= "html")

# Create a list of models
tot_m <- dlply(chasamil_tot, .(Country), dynModelCountry)
tot_m.coefs <- ldply(tot_m, country_series_table)

```

```{r results='asis'}
# Print table to html
print(xtable(tot_m.coefs,
             caption="Demand equations for total paper and paperboard by country", 
             label="DemandByCountry"),  type = "html")

```


### Same model from 1978 only
```{r results='asis'}
# Model with GDP and Price in constant USD from 1978
tot_m <- dlply(subset(chasamil_tot, Year>=1978), .(Country), dynModelCountry)
tot_m.coefs <- ldply(tot_m, country_series_table)
print(xtable(tot_m.coefs,
             caption="Tot paper Demand Model with GDP and Price in constant USD from 1978", 
             label="DemandByCountry"),  type = "html")
```

### Same model from 1978 only using EUR price and GDP values
```{r results='asis'}
# Model with GDP and Price in constant EUR from 1978
dirty <- subset(chasamil_tot, Year>=1978)
# Rename GDPconstantEUR to GDPconstantUSD to avoid creating new functions 
#     for dynModelCountry() and country_series_table()
dirty$GDPconstantUSD <- dirty$GDPconstantEUR
tot_m_EUR <- dlply(dirty, .(Country), dynModelCountry)
tot_m_EUR.coefs <- ldply(tot_m_EUR, country_series_table)
print(xtable(tot_m_EUR.coefs,
             caption="Tot paper Demand Model with GDP and Price in constant EUR from 1978", 
             label="DemandByCountry"),  type = "html")



```

On the subject of Cochrane Orcutt method, [John Fox wrote](https://stat.ethz.ch/pipermail/r-help/2002-January/017774.html): "I'm not sure why you'd want to use these methods, other than for 
historical reasons." And he gave a function cochrane.orcutt.lm which takes a linear-model object as an input.


Autocorrelation
---------------
```{r eval=FALSE}
fr <- subset(chasamil, Country=="France", select=c("Consumption","GDPconstantUSD",
                                                   "Price", "Consum_t_1"))

m_fr = lm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1),
          data=fr)
summary(m_fr)

resid <- data.frame(r = m_fr$residuals[-1],
                    r_t_1 = m_fr$residuals[-length(m_fr$residuals)])

mr_fr <- lm(r ~ r_t_1, data=resid)
summary(mr_fr)


# Function from https://stat.ethz.ch/pipermail/r-help/2002-January/017774.html
cochrane.orcutt.lm <- function(mod){
     X <- model.matrix(mod)
     y <- model.response(model.frame(mod))
     e <- residuals(mod)
     n <- length(e)
     names <- colnames(X)
     rho <- sum(e[1:(n-1)]*e[2:n])/sum(e^2)
     y <- y[2:n] - rho * y[1:(n-1)]
     X <- X[2:n,] - rho * X[1:(n-1),]
     mod <- lm(y ~ X - 1)
     result <- list()
     result$coefficients <- coef(mod)
     names(result$coefficients) <- names
     summary <- summary(mod, corr = F)
     result$cov <- (summary$sigma^2) * summary$cov.unscaled
     dimnames(result$cov) <- list(names, names)
     result$sigma <- summary$sigma
     result$rho <- rho
     class(result) <- 'cochrane.orcutt'
     result
     }

summary(m_fr)
cochrane.orcutt.lm(m_fr)

# Plot autocorelation of residuals
e <- m_fr$residuals
n <- length(e)
plot(e[1:(n-1)],e[2:n])

# Same model for Fnland
m_fi = lm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1),
           data=subset(chasamil, Country=="Finland", 
                       select=c("Consumption","GDPconstantUSD","Price", "Consum_t_1")))
summary(m_fi)
cochrane.orcutt.lm(m_fi)


m_au = lm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1),
           data=subset(chasamil, Country=="Austria", 
                       select=c("Consumption","GDPconstantUSD","Price", "Consum_t_1")))
summary(m_au)
cochrane.orcutt.lm(m_au)
```


Correlation between model variables
```{r eval=FALSE}
cor(fr)
```



Statistical tests
-----------------

### Serial temporal autocorrelation on one country
```{r}
# Change to a time series
consfr <- subset(chasamil,Country=="France")
consfr.ts <- ts(consfr$Consumption, start=min(consfr$Year))
plot(consfr.ts)
acf(consfr.ts,lag.max=100)
# numerical values of this autocorrelation function
a <- acf(consfr.ts,lag.max=100,plot=FALSE)
class(a)


# Autocorrelation coefficients to be applied on all countries
autocorr <- function(dtf){
    cons.ts <- ts(dtf$Consumption, start=min(dtf$Year))
    return(acf(cons.ts,lag.max=100,plot=FALSE)$acf)
}
a = ddply(chasamil, .(Country), autocorr)
autocorrTable <- melt(a, id.vars = c("Country"))
ggplot(autocorrTable) + aes(x=variable, y=value) +
    geom_bar(stat="identity") + facet_wrap(~Country)
```



Panel data models
-----------------
Over the 1960 - 1992 estimation period.
```{r}
# Set data as panel data
pchasamil <- plm.data(chasamil_tot, index=c("Country","Year"))
```

### Pooled OLS
Estimation with pooled OLS
$$ln D_it = \delta_0 + \delta_1 ln Y + \delta_2 ln P + \delta_3 ln D_-1 + \epsilon$$

Oneway (individual) effect Pooling Model
```{r, results='asis'}
pooling <- plm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1), 
               data=pchasamil, model= "pooling")
printSummaryTable(pooling)
```

* with the package plm 
```{r}
summary(pooling)
```

* With base lm is it the same model?
```{r}
pooling2 <- plm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1), 
               data=chasamil_tot)
summary(pooling2)

# How about adding individual country effects?
pooling3 <- plm(log(Consumption) ~ Country + log(GDPconstantUSD) + 
                    log(Price) + log(Consum_t_1), 
               data=chasamil_tot)
summary(pooling3)
```
```
                         Estimate Std. Error t-value  Pr(>|t|)   
* log(GDPconstantUSD)    0.0758327  0.0205783  3.6851 0.0002759 
* log(Price)            -0.0330761  0.0607351 -0.5446 0.5864781    
* log(Consum_t_1)        0.5330548  0.0508572 10.4814 < 2.2e-16 
```
### Fixed effects or within estimator
```{r, results='asis'}
fixed <- plm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1), 
               data=pchasamil, model= "within")
printSummaryTable(fixed)
```

### Random effects estimator
```{r, results='asis'}
random <- plm(log(Consumption) ~ log(GDPconstantUSD) + log(Price) + log(Consum_t_1), 
               data=pchasamil, model= "random")
printSummaryTable(random)
```
The estimation of the random effect model tells me that there is an Error in swar(object, data, effect) : "the estimated variance of the individual effect is negative".

